---
title: "R Exercises 4.4 - Data Import/Export"
author: "Student Name"
date: "`r Sys.Date()`"
output: html_document
---

## Exercise 1
Read in the Youth Tobacco study data and name it youth

```{r}
# Method 1: Try with UTF-8 encoding
youth <- read.csv("Youth_Tobacco_Survey__YTS__Data.csv", 
                  header = TRUE, 
                  encoding = "UTF-8")

# Alternative methods if above doesn't work:
# Method 2: Try with latin1 encoding
# youth <- read.csv("Youth_Tobacco_Survey__YTS__Data.csv", 
#                   header = TRUE, 
#                   encoding = "latin1")

# Method 3: Use readr package
# library(readr)
# youth <- read_csv("Youth_Tobacco_Survey__YTS__Data.csv", 
#                   locale = locale(encoding = "UTF-8"))

head(youth, 3)
class(youth)
dim(youth)
```

## Exercise 2
Install and invoke the readxl package

```{r}
install.packages("readxl")
library(readxl)
```

## Exercise 3
Download an Excel version of the Superstore dataset from Blackboard. Use the read_excel() function in the readxl package to read in the dataset and call the output super.

```{r}
super <- read_excel("Superstore.xlsx")
```

### 3a. What types of variables are in the dataset? (numeric, text, yes/no)

```{r}
str(super)
sapply(super, class)
```

### 3b. How many rows and columns are there?

```{r}
nrow(super)
ncol(super)
dim(super)
```

### 3c. What are the basic statistics for each variable? (mean, min, max, counts)

```{r}
summary(super)
```

### 3d. Are there any missing values?

```{r}
sum(is.na(super))
colSums(is.na(super))
```

### 3e. How many unique values are in each variable?

```{r}
sapply(super, function(x) length(unique(x)))
```

## Exercise 4
Write out the super R object as a CSV file using write.csv() and call the file "superstore.csv"

```{r}
write.csv(super, file = "superstore.csv", row.names = FALSE)
```