---
title: "R Exercises 6.8 - Subsetting Data"
author: "Student Name"
date: "`r Sys.Date()`"
output: html_document
---

## Exercise 1
Check if the mtcars dataset is available by typing mtcars.

```{r}
mtcars
```

## Exercise 2
What is the class of mtcars?

```{r}
class(mtcars)
```

## Exercise 3
How many rows (observations) and columns (variables) does it have?

```{r}
nrow(mtcars)
ncol(mtcars)
dim(mtcars)
```

## Exercise 4
Make a copy of mtcars called cars, then rename the column mpg to MPG.

```{r}
cars <- mtcars
colnames(cars)[1] <- "MPG"
head(cars)
```

## Exercise 5
Make a new dataset carsSub containing just the columns wt, qsec, and hp, then check its dimensions with dim().

```{r}
carsSub <- cars[, c("wt", "qsec", "hp")]
dim(carsSub)
head(carsSub)
```

## Exercise 6
Filter the rows of cars where mpg > 20. How many rows are returned?

```{r}
# Note: using MPG since we renamed the column
filtered_cars <- cars[cars$MPG > 20, ]
nrow(filtered_cars)
filtered_cars
```

## Exercise 7
Filter the rows where mpg < 16 and hp > 100. How many rows are returned?

```{r}
# Note: using MPG since we renamed the column
filtered_cars2 <- cars[cars$MPG < 16 & cars$hp > 100, ]
nrow(filtered_cars2)
filtered_cars2
```

## Exercise 8
Create a new carsSub that contains only the columns wt, qsec, and hp for cars with 8 cylinders (cyl). What are its dimensions?

```{r}
carsSub <- cars[cars$cyl == 8, c("wt", "qsec", "hp")]
dim(carsSub)
carsSub
```

## Exercise 9
Reorder carsSub by wt in ascending order (use arrange()).

```{r}
library(dplyr)
carsSub <- arrange(carsSub, wt)
carsSub
```

## Exercise 10
Add a new variable wt2 to carsSub equal to wt^2.

```{r}
carsSub$wt2 <- carsSub$wt^2
carsSub
```