---
title: "R Exercises 7.3 - Data Summarization"
author: "Student Name"
date: "`r Sys.Date()`"
output: html_document
---

**Note:** These exercises use the Bike Lanes.csv dataset. Since the actual file is not available in this workspace, the code is provided but will need the actual dataset to run.

## Exercise 1
How many bike lanes are in Baltimore? (Each row = one lane.)

```{r eval=FALSE}
# Read the dataset
bike_lanes <- read.csv("Bike Lanes.csv")

# Count the number of rows (bike lanes)
nrow(bike_lanes)
```

## Exercise 2
How much total bike lane length is in Baltimore in (a) feet and (b) miles? (length is given in feet.)

```{r eval=FALSE}
# (a) Total length in feet
total_feet <- sum(bike_lanes$length, na.rm = TRUE)
total_feet

# (b) Total length in miles (1 mile = 5280 feet)
total_miles <- total_feet / 5280
total_miles
```

## Exercise 3
How many bike lane types (type) are there? Which type is (a) most common and (b) longest on average?

```{r eval=FALSE}
# Number of bike lane types
length(unique(bike_lanes$type))

# Count of each type
type_counts <- table(bike_lanes$type)
type_counts

# (a) Most common type
names(which.max(type_counts))

# (b) Average length by type
library(dplyr)
type_summary <- bike_lanes %>%
  group_by(type) %>%
  summarise(avg_length = mean(length, na.rm = TRUE)) %>%
  arrange(desc(avg_length))

type_summary
# Longest on average
type_summary$type[1]
```

## Exercise 4
How many projects (project) are there, and which one has the longest average bike lane length?

```{r eval=FALSE}
# Number of projects
length(unique(bike_lanes$project))

# Average length by project
project_summary <- bike_lanes %>%
  group_by(project) %>%
  summarise(avg_length = mean(length, na.rm = TRUE)) %>%
  arrange(desc(avg_length))

project_summary
# Project with longest average length
project_summary$project[1]
```

## Exercise 5
What is the average bike lane length per year of installation? (Set dateInstalled = NA if it is 0.)

```{r eval=FALSE}
# Set dateInstalled = NA if it is 0
bike_lanes$dateInstalled[bike_lanes$dateInstalled == 0] <- NA

# Average length by year of installation
year_summary <- bike_lanes %>%
  filter(!is.na(dateInstalled)) %>%
  group_by(dateInstalled) %>%
  summarise(avg_length = mean(length, na.rm = TRUE)) %>%
  arrange(dateInstalled)

year_summary
```

## Exercise 6
Describe bike lane lengths (a) with numbers and (b) with a plot.

```{r eval=FALSE}
# (a) Numerical summary
summary(bike_lanes$length)
mean(bike_lanes$length, na.rm = TRUE)
sd(bike_lanes$length, na.rm = TRUE)
median(bike_lanes$length, na.rm = TRUE)
quantile(bike_lanes$length, na.rm = TRUE)

# (b) Plot
hist(bike_lanes$length, 
     main = "Distribution of Bike Lane Lengths",
     xlab = "Length (feet)",
     breaks = 20)

boxplot(bike_lanes$length,
        main = "Boxplot of Bike Lane Lengths",
        ylab = "Length (feet)")
```

## Exercise 7
Describe lane lengths again, but grouped by (a) type and (b) number of lanes (numLanes).

```{r eval=FALSE}
# (a) By type
library(dplyr)
bike_lanes %>%
  group_by(type) %>%
  summarise(
    count = n(),
    mean_length = mean(length, na.rm = TRUE),
    median_length = median(length, na.rm = TRUE),
    sd_length = sd(length, na.rm = TRUE)
  )

# Boxplot by type
boxplot(length ~ type, data = bike_lanes,
        main = "Bike Lane Lengths by Type",
        xlab = "Type",
        ylab = "Length (feet)")

# (b) By number of lanes
bike_lanes %>%
  group_by(numLanes) %>%
  summarise(
    count = n(),
    mean_length = mean(length, na.rm = TRUE),
    median_length = median(length, na.rm = TRUE),
    sd_length = sd(length, na.rm = TRUE)
  )

# Boxplot by number of lanes
boxplot(length ~ numLanes, data = bike_lanes,
        main = "Bike Lane Lengths by Number of Lanes",
        xlab = "Number of Lanes",
        ylab = "Length (feet)")
```