---
title: "R Exercises 3.4 - Basic R Operations"
author: "Student Name"
date: "`r Sys.Date()`"
output: html_document
---

## Exercise 1
Create a new variable called my.num that contains 6 numbers of your choice

```{r}
my.num <- c(2, 5, 8, 12, 15, 20)
my.num
```

## Exercise 2
Multiply my.num by 4

```{r}
my.num * 4
```

## Exercise 3
Create a second variable called my.char that contains 5 character strings of your choices

```{r}
my.char <- c("apple", "banana", "cherry", "date", "elderberry")
my.char
```

## Exercise 4
Combine the two variables my.num and my.char into a variable called both

```{r}
both <- c(my.num, my.char)
both
```

## Exercise 5
What is the length of both?

```{r}
length(both)
```

## Exercise 6
What class is both?

```{r}
class(both)
```

## Exercise 7
Divide both by 3, what happens?

```{r}
both / 3
```

## Exercise 8
Create a vector with elements 1 2 3 4 5 6 and call it x

```{r}
x <- c(1, 2, 3, 4, 5, 6)
x
```

## Exercise 9
Create another vector with elements 10 20 30 40 50 and call it y

```{r}
y <- c(10, 20, 30, 40, 50)
y
```

## Exercise 10
What happens if you try to add x and y together? why?

```{r}
x + y
```

## Exercise 11
Append the value 60 onto the vector y (hint: you can use the c() function)

```{r}
y <- c(y, 60)
y
```

## Exercise 12
Add x and y together

```{r}
x + y
```

## Exercise 13
Multiply x and y together. Pay attention to how R performs operations on vectors of the same length.

```{r}
x * y
```